---
title: "Thermocouple Notes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r,include=FALSE}
#install.packages("tidyverse")
#install.packages("gridExtra")
#install.packages("here")
library(here)
library(tidyverse)
library(ggplot2)
library(gridExtra)
setwd(here())
data<-c('Therm_K fit.csv','HallFermium2.csv')
```

**THIS DOCUMENT IS ONLY A DRAFT**

# Thermocopule temperature fit

```{r}
therm.K<-read.csv(here('analysis/data',data[1]),sep=",")
```

```{r,include=FALSE}
colnames(therm.K)<-c("T","Ek","CalcT4","Error4","CalcT5","Err5")
attach(therm.K)
```

```{r}
TEkPlot<-ggplot(data=therm.K,aes(x=Ek,y=T))+
         geom_line()+
         ggtitle('Thermocopule value vs Temperature')+
         theme(plot.title = element_text(hjust = 0.5))
TEkPlot
```

This document contains a few notes that explains the math behind the polynomial used to convert from the raw E of the termocouple to the temperature. The official [NIST termocouple tables](https://srdata.nist.gov/its90/download/download.html) are not provided to cover for the +-150°c temperature range the apparatus uses. For this reason, it was necessary to do some trivial fitting to recalculate the polynomials.

A typical thermocouple conversion polynomial is as follow:

\begin{equation}
t_{\text{calc}}=d_0+d_1 E+d_2 E^2+\ldots + d_n E^n
\end{equation}

A fitting polynomial of the fifth order is sufficient, given the precision of our equipment.
Best fit coefficients obtained from NIST data tables ($-200 < t [^{\circ}C]<200$)

| Coefficients | Value |
:--------------|:------|
$d_0$ | 0.3837 |
$d_1$ | 25.22 |
$d_2$ | 0.2795 |
$d_3$ | 0.07205 |
$d_4$ | 0.01409 |
$d_5$ | 0.001056 |

\begin{equation}
T = \beta_0+\beta_1 E + \ldots + \beta_5 E^5
\end{equation}
